---
output:
  word_document:
    reference_docx: cntyReport_styles.docx
always_allow_html: yes
params:
  c: NA
---

```{r setup, include=FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


```{r process, echo=FALSE, warning = FALSE, message=TRUE }
#Packages.
#application
require(rmarkdown)
require(knitr)
require(shiny)
#data
library(dplyr)
#library(fs)
library(readr)
library(ggplot2)
library(shinyBS)
library(magrittr)
library(tidyr)
library(plotly)
library(shinyjs)
library(gridGraphics)

```
###`r paste0(' Measures of Health Status Snapshot',": ",DataYearBegin," to ",DataYearEnd)`
**Report downloaded `r paste0(format(today, format="%m-%d-%Y"))`**

Public health looks across multiple measures to identify significant patterns and trends and public health challenges.

```{r echo=FALSE, fig.height=7, fig.width=7, warning=FALSE, , message=TRUE, dev='CairoPNG'}
# Page 1, Sub-Layout

P1.Plot1 <- renderPlot({plotMeasures(IDnum=1,input$myCounty)})
P1.Plot2 <- renderPlot({plotMeasures(IDnum=2,input$myCounty)})
P1.Plot3 <- renderPlot({plotMeasures(IDnum=3,input$myCounty)})
P1.Plot4 <- renderPlot({plotMeasures(IDnum=4,input$myCounty)})

left_col  <- plot_grid(P1.Plot1, P1.Plot2, P1.Plot3, P1.Plot4, labels = NULL, ncol = 1, align = 'v')
right_col <- plot_grid(P1.Map1, P1.Bar1, labels = NULL, ncol = 1, rel_heights = c(2, 1))

```


```{r echo=FALSE, fig.height=7, fig.width=7, warning=FALSE, , message=TRUE, dev='CairoPNG'}
# Page 1, Layout
op <- par(omi=c(0,0,0,0))
plot_grid(left_col, right_col, labels = NULL, ncol = 2, rel_widths= c(1,1))
par(op)
```

###Overall State Observations

Number of Deaths
While ischemic heart disease has decreased, it remains a top ranked condition based on multiple measures including total number of deaths. Alzheimer’s disease was the second ranked condition.

Premature Deaths
The first and third highest ranked conditions are unintentional injury and substance use, two conditions affecting relatively younger people. Ischemic heart disease is ranked third because of the large numbers of deaths from this condition.

Greatest Percentage Increase in Deaths
Deaths from several conditions have increased substantially in California from 2008 to 2018 including Kidney diseases, Alzheimer’s disease, and Congestive heart failure.

Disparity Ratio
Persistent disparities in health remain, with a range of social determinants of health, leading to some populations experiencing higher burdens of disease than others. The widest gaps are for HIV/AIDS, Alcohol use disorders, Opioid use disorders, and Homicide, in which the rates for the group with the worst outcomes are many times higher (16.4 to 12.6) than the group with the best outcomes.
